---
title: "IS 609 Week 11 Homework"
author: "Ben Arancibia"
date: "November 5, 2015"
output: pdf_document
---

__529__

In Problems 5-8, find and classify the rest points of the given autonomous system.

6)

dx/dt = -(y-1)

dy/dt = x-2

The rest point is (2,1). Use deSolve package to help us visualize the various trajectories.

```{r, echo=FALSE}
library(deSolve)
library(ggplot2)
fn <- function(t, yv, b) 
{
  x <- yv[1]
  y <- yv[2]
  
  x.p <- (-1 * (y - 1))
  y.p <- (x-2)
  
  return(list(c(x=x.p,y=y.p)))
}

h <- 0.1
t <- seq(0, 10, h)
id <- 0
dfRk <- data.frame(id=c(), x=c(), y=c())
for(x in seq(0, 3, 0.5))
{
  for(y in seq(0, 3, 0.5))
  {
    y0 <- c(x,y)
    id <- id + 1
    rk45 <- rk(y0, t, func=fn, parms=1, method="rk45dp7")
    dfRk <- rbind(dfRk, cbind(id=id, x=rk45[, 2], y=rk45[,3]))
  }
}

head(dfRk, 10)
```

Based on the visualization, (2,1) is stable rest point.

```{r}
g1 <- ggplot(data=dfRk) +
  geom_path(aes(x=x, y=y, colour=id, group=id)) 
g1

```


__536__

7) Show that the two trajectories leading to m/n, a/b shown in Figure 12.8 are unique.

a)

\[\frac{dy}{dx}=\frac{(m-nx)y}{(a-b)x}\]

Dividing two autonomous differential equations with each other produces the derivative \(\frac{dy}{dx}\)

\[\frac{dy}{dx}=\frac{\frac{dy}{dt}}{\frac{dx}{dr}}=\frac{(m-nx)y}{(a-b)x}\]

b)

\[\frac{dy}{dx}=\frac{(m-nx)y}{(a-b)x}\]

\[\frac{(a-b) dy}{y}=\frac{(m-nx) dx}{x}\]

\[\frac{a}{y} - \frac{b}{y} dy = \frac{m}{x} - n dx\]

Integrate both sides:

\[a \text{ ln } y - b \text{ ln } y = m \text{ ln } x - nx + K\]

\[a \text{ ln } y - b \text{ ln } y - m \text{ ln } x + nx = K\]

Simplifying, we get:

\[y^a e^{-by} = Kx^me^{-nx}\]

c)
d)
e)
f)

__546__

1) Apply the first and second derivative tests to the function f(y) = y^a / e^by to show that y=a/b is a unique critical point that yields the relative maximum f(a/b) Show also that f(y) approaches zero as y tends to infinity.


First derivative:
f(y) = y^a/e^by 

f(y) = y^a * e^-by

product rule:

f'(y) = y^a * -b e^-by + ay^a-1 * e^-by

Sove for y

f'(y) = y^a * -b e^-by + ay^a-1 * e^-by = 0

f'(y) = e^-by (-by^a + ay^a-1) = 0

2 zeroes of the first derivative. 

e^-by = 0  and  -by^a + ay^a-1 = 0

Solve for y from equation -by^a + ay^(a-1) = 0:

-by^a + ay^a-1 = 0

ay^(a-1) = by^a

a = by^a/y^(a-1) = by

(a)/(b) = y

Determining second derivative:

f'(y) = e^(-by) (-by^a + ay^(a-1)

product rule:


