---
title: "IS 609 Week 13 Homework"
author: "Ben Arancibia"
date: "November 17, 2015"
output:
  pdf_document: default
---

__B-13__

4) Discuss how you might go about validating the nuclear arms race model. What data would you collect? Is it possible to obtain the data?

One way that we go about validating the nuclear arms race model is to look at historical records. There are tables out there with inventories, but it is difficult to determine how accurate these numbers are and usually records are not up to date. The National Resources Defense Council has table of 1945 - 2002 for nuclear weapons inventories: http://www.nrdc.org/nuclear/nudb/datab19.asp and the Bullentin of Atomic Scientists also has a table of inventories from 1945- 2010: http://bos.sagepub.com/content/66/4/77.full.pdf+html. With this data you could then estimate the survivability of each country's missle silos. Once you have this data you can then create standard deviations and moving averages to give a range of what the inventories might be for each country. 


__B-17__

1) Build a numercial solution to Equations (15.8)

yn+1 = 120 + 1/2 * xn

xn+1 = 60 + 1/3 * yn

with

[x_0=100]

[y_0=200]

a) Graph your results

```{r}

library(ggplot2)
model <- function(x.naught, y.naught)
{
  xn <- x.naught
  yn <- y.naught
  model.results <- data.frame(i=0,xn=c(xn), yn=c(yn))
  for(i in 1:20)
  {
    yn1 <- 120 + (1/2 * xn)
    xn1 <- 60 + (1/3 * yn)
    
    xn <- xn1
    yn <- yn1
    
    model.results <- rbind(model.results, cbind(i=i, xn=xn, yn=yn))
  }
  
  return (model.results)
}
# Run the model and get the resulting cycle data
nuke.race<- model(100, 200)

g1 <- ggplot(data=nuke.race) + geom_line(aes(x=i, y=xn), colour="green") + 
  geom_line(aes(x=i, y=yn), colour="blue")

g1
```

b) Is an equilibrium value reached

Yes equilibirum is reached.

c) Try other starting values. Do you think the equlibrium is stable

```{r}
nuke.race<- model(0, 0)

g1 <- ggplot(data=nuke.race) + geom_line(aes(x=i, y=xn), colour="green") + 
  geom_line(aes(x=i, y=yn), colour="blue")

g1

nuke.race<- model(1000, 1000)

g1 <- ggplot(data=nuke.race) + geom_line(aes(x=i, y=xn), colour="green") + 
  geom_line(aes(x=i, y=yn), colour="blue")

g1

```

Yes equilibrium is stable
d) Explore other values for survival coefficients of Countries X and Y. Describe your results.

```{r}
model <- function(x.naught, y.naught)
{
  xn <- x.naught
  yn <- y.naught
  model.results <- data.frame(i=0,xn=c(xn), yn=c(yn))
  for(i in 1:20)
  {
    yn1 <- 120 + (1/8 * xn)
    xn1 <- 60 + (2/3 * yn)
    
    xn <- xn1
    yn <- yn1
    
    model.results <- rbind(model.results, cbind(i=i, xn=xn, yn=yn))
  }
  
  return (model.results)
}
# Run the model and get the resulting cycle data
nuke.race<- model(100, 200)

g1 <- ggplot(data=nuke.race) + geom_line(aes(x=i, y=xn), colour="green") + 
  geom_line(aes(x=i, y=yn), colour="blue")

g1

```

Changing survival coeffcients changes which nation will end up with largest inventory of nuclear weapons. 
